---
title: "Hisensitivity-CRP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(knitr)
library(gtools)
```

## R Markdown

Extract the peak value of hisensitivity c-reactive protein for each patient in EDW cerner table. Convert those who have mg/dL as measure unit to mg/l. Visualize the CRP value using histogram and boxplot.  

```{r gtools}{r plyr}
setwd('/Volumes/fsmhome/projects/new_dataset_peakval_only/hisensitivity c-reactive protein/data/')
x=read.csv('./diagnosis-hisencrp_pk.csv',header = 0 )
colnames(x) = c('mrd_pt_id','event_cd','vocabulary_value','result_val_num','result_units_dsc','event_start_dt_tm','diagnosis_dt')
#convert between units if it is high sensitivity reactive protein 
x[x$result_units_dsc=='mg/dL','result_val_num'] = x[x$result_units_dsc=='mg/dL','result_val_num']*10
x$val_disc= as.factor(as.numeric(quantcut(x$result_val_num,3)))
summary(x$result_val_num)
```

```{r pressure, echo=FALSE}
title = 'Hisensitivity CRP Peak Histogram'
xlab= 'Hisensitivity CRP value (mg/L)'
ggplot(data=x , aes(result_val_num)) + geom_histogram(binwidth = 3,fill='darkorange',col='black') +
  labs(title=title, x=xlab)+ theme(axis.text = element_text(size = 15)
   ,axis.title = element_text(size=18), plot.title = element_text(size=20,hjust=0.5)) 

title = ''
xlab= 'Hisensitivity CRP'
ylab = 'Hisensitivity CRP Value  (mg/L)'
ggplot(x, aes(x = '', y = result_val_num)) +
  geom_boxplot(fill = 'darkorange', colour = 'black')  + 
  labs(title=title, x=xlab, y=ylab)+ theme(axis.text = element_text(size = 15)
  ,axis.title = element_text(size=18), plot.title = element_text(size=20,hjust=0.5))
```
Create a factor variable using the quantiles of a continous variable.
```{r}
summary(quantcut(x$result_val_num,3))
```
Run chi-square test for each diagnosis code. Compare the odds ratio between group 2 and group1, group3 and group1. 
```{r code, eval = FALSE}

#focus on diagnoses that have frequency higher than 10. 
freq = as.data.frame(table(x$vocabulary_val))
#freq=freq[freq$Var1!='*****',]
dx = freq[freq$Freq>=10,]  #change threshold as you want 
dx_list = as.character(dx$Var1)
result =  data.frame()
i = 1
for (voca in dx_list)
{ 
  # voca = '250.00'
  df_case = x[x$vocabulary_val==voca,]  #select rows with dx voca 
  case_id = unique(df_case$mrd_pt_id) #find all the cases id 
  label = as.data.frame(cbind(case_id, rep('case',length(case_id))))
  colnames(label) = c('case_id', 'label')
  #control_id =  setdiff(unique(x$mrd_pt_id), unique(df_case$mrd_pt_id)) #find all the control id 
  #df_control = x[x$mrd_pt_id %in% control_id, ] # create control dataset 
  df_lab = merge(x,label, by.x='mrd_pt_id',by.y='case_id',all.x=TRUE)
  df_lab$label = as.character(df_lab$label)
  df_lab$label[is.na(df_lab$label)] <- 'ctrl'
  tb = table(df_lab$label, df_lab$val_disc) 
  res12= fisher.test(tb[,1:2]) #chi-sqaure test for more than 2 categories 
  res13= fisher.test(tb[,c(1,3)])
  result[i,'diagnosis'] = voca
  result[i, 'OR_1vs2'] = res12$estimate
  result[i, 'pval_1vs2'] = res12[1]
  result[i, 'conf.int_1v2']=paste('[',res12$conf.int[1],',',res12$conf.int[2],']')
  result[i, 'OR_1vs3'] = res13$estimate
  result[i, 'pval_1vs3'] = res13[1]
  result[i, 'conf.int_1v3']=paste('[',res13$conf.int[1],',',res13$conf.int[2],']')
  i = i+ 1 
  print(i)
  
}
```
Read the ICD9 dictionary to get the description of each ICD9 code. 
```{r}
dict = read.csv('/Volumes/fsmhome/projects/new_dataset_peakval_only/hisensitivity c-reactive protein/data/icd-hisenCRP-after.txt',sep = '|',header= 0)
dict = dict[!duplicated(dict$V1),]
dict = dict[dict$V2!='',]
final = merge(result, dict, by.x= 'diagnosis', by.y='V1',all.x=T)
final = final[,c('diagnosis','V2','OR_1vs2','pval_1vs2','conf.int_1v2','OR_1vs3','pval_1vs3','conf.int_1v3')]
colnames(final)[2]='icd9_dsc'
final1 = merge(final,freq, by.x= 'diagnosis', by.y= 'Var1',all.x=T)
```

```{r}
#result = read.csv('/Volumes/fsmhome/projects/new_dataset_peakval_only/hisensitivity c-reactive protein/result/hisenCRP_diagnosis_chisquare.csv')
kable(final1[order(final1$Freq, decreasing = T),][1:200,])
```
List those who are siginificant unique to group2/1 but not to group3/1
```{r, echo= F}
data = final1 
group2vs1= data[data$pval_1vs2<=0.05 & data$pval_1vs3>0.05, ]
dim(group2vs1)
kable(group2vs1[order(group2vs1$Freq, decreasing = T),])
```
List those who are significant unique to group3/1 but not 2/1 
```{r}
group3vs1= data[data$pval_1vs2>0.05 & data$pval_1vs3<=0.05, ]
dim(group3vs1)
kable(group3vs1[order(group3vs1$Freq, decreasing = T),])

```
List those who are significant in both groups 
```{r}
group23= data[data$pval_1vs2<=0.05 & data$pval_1vs3<=0.05, ]
dim(group23)
kable(group23[order(group23$Freq, decreasing = T),])
```

```{r, echo=FALSE}
sample= x[x$vocabulary_val =='723.1',]
d=ggplot() + geom_density(data= sample, aes(x=result_val_num, fill='ICD723.1:Neck pain',alpha=0.5))+ xlab('hisen-CRP value')+ylab('density') + coord_cartesian(xlim=c(0, 50))
d+ geom_density(data= x, aes(x=result_val_num, fill='population', alpha = 0.5)) 
```
```{r, eval=FALSE}
voca = '723.1'
df_case = x[x$vocabulary_val==voca,]  #select rows with dx voca 
case_id = unique(df_case$mrd_pt_id) #find all the cases id 
label = as.data.frame(cbind(case_id, rep('case',length(case_id))))
colnames(label) = c('case_id', 'label')
#control_id =  setdiff(unique(x$mrd_pt_id), unique(df_case$mrd_pt_id)) #find all the control id 
#df_control = x[x$mrd_pt_id %in% control_id, ] # create control dataset 
df_lab = merge(x,label, by.x='mrd_pt_id',by.y='case_id',all.x=TRUE)
df_lab$label = as.character(df_lab$label)
df_lab$label[is.na(df_lab$label)] <- 'ctrl'
#tb = table(df_lab$label, df_lab$val_disc) 
#rownames(tb) = c(paste('icd',voca),'control')
kable(tb)
```



