---
title: "scatter plot_hisenCRP"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(gtools)
library(ggplot2)
library(knitr)
library(reshape2)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Read chi-square and diagnosis association file(file can be found here: https://northwestern.app.box.com/folder/46902279768). This file has information of pvalue of chi-square association and oddsRatio. Hisensitivty CRP values are randomly cut into 8 categories, with category 1: 0-1mg/l being normal, and the following:  1-3,3-5,5-10,10-20,20-30,30-50,50-100, >100
```{r, eval = F}
x$val_disc = as.numeric(cut(x$result_val_num, c(0,1,3,5,10,20,30,50,100)))
```

```{r cars}
data = read.csv('/Users/yudeng/name')
data$dx_str = data$diagnosis
data$diagnosis = as.character(data$diagnosis)
df_p = data[,c('dx_str','pval_12','pval_13','pval_14','pval_15',
               'pval_16','pval_17')]
df1 = melt(df_p, value.name='p_val')
df_or = data[,c('dx_str','OR_12','OR_13','OR_14','OR_15','OR_16','OR_17')]
df2= melt(df_or, value.name='OddsRatio')

dd = cbind(df1,df2)
dd= dd[,c('dx_str','variable','p_val',"OddsRatio")]
dd_plot = dd[dd$p_val< 0.05, ]
## get diagnosis-hisCRP matrix plot 

dd_plot$OddsRatio = as.character(dd_plot$OddsRatio)
dd_plot$OddsRatio = as.numeric(dd_plot$OddsRatio )
dd_plot$indictor =as.numeric(dd_plot$OddsRatio < 1)
dd_plot[dd_plot$indictor==0,'indictor']='decrease_risk'
dd_plot[dd_plot$indictor==1,'indictor']='increase_risk'

###dd_plot1 dropped icd with Inf value which means 0 cell counts in that category 
dd_plot1 = dd_plot
dd_plot1 = dd_plot1[dd_plot1$OddsRatio!='Inf',]
head(dd_plot1)

###dd_plot include inf value and mark it with a different indictor which will have a different color mark 
dd_plot[dd_plot$OddsRatio=='Inf','indictor'] = 'Inf'
dd_plot= dd_plot[order(dd_plot$dx_str),]
dd_plot$dx_str <- factor(dd_plot$dx_str, levels=unique(dd_plot$dx_str))

dd_plot1= dd_plot1[order(dd_plot1$dx_str),]
dd_plot1$dx_str <- factor(dd_plot1$dx_str, levels=unique(dd_plot1$dx_str))

```

## Including Plots
The dots represent 1/odds ratio, the larger the dots, the higher the risk of getting certain diagnosis compared to normal lab range (hisensitivity 0-1). In the category where the lab value is high, some cell counts go down to zero, oddsratio becomes inf, we use a different color to mark those with OR of inf. 
```{r pressure, echo=FALSE,fig.width=7,fig.height=7}
##create p1, p2, which include and exclude 'Inf' value and add color 
p1=ggplot(dd_plot[1:300,], aes(x = variable, y = as.factor(dx_str),color=as.factor(indictor))) + geom_point(aes(size = (as.numeric(1/OddsRatio)))) + scale_size_continuous(range = c(0.5, 3))
p2=ggplot(dd_plot1[1:300,], aes(x = variable, y = as.factor(dx_str),color=as.factor(indictor))) + geom_point(aes(size = (as.numeric(1/OddsRatio)))) + scale_size_continuous(range = c(0.5, 3))
#grid.arrange(p1, p2, nrow = 1)

###repeat to see different rows in the dataframe 
p3=ggplot(dd_plot[300:600,], aes(x = variable, y = as.factor(dx_str),color=as.factor(indictor))) + geom_point(aes(size = (as.numeric(1/OddsRatio)))) + scale_size_continuous(range = c(0.5, 3))
p4=ggplot(dd_plot1[300:600,], aes(x = variable, y = as.factor(dx_str),color=as.factor(indictor))) + geom_point(aes(size = (as.numeric(1/OddsRatio)))) + scale_size_continuous(range = c(0.5, 3))
p1
p2
##arrange p1 and p2 and p3 and p4 
#grid.arrange(p1, p2, p3, p4, nrow = 4)
```


